\documentclass[10pt,twocolumn]{article}

% ── Packages ──────────────────────────────────────────────────────────
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{times}
\usepackage[margin=0.75in,top=0.7in,bottom=0.7in,columnsep=0.3in]{geometry}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{microtype}
\usepackage{amsmath,amssymb}
\usepackage{caption}
\usepackage{url}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,shapes.geometric,fit,calc}

% ── Hyperlinks ────────────────────────────────────────────────────────
\hypersetup{
  colorlinks=false,
  pdfborder={0 0 0},
  pdftitle={DeltaScribe Edge: Offline Longitudinal CXR Copilot},
  pdfauthor={Simon},
}

% ── Section formatting ────────────────────────────────────────────────
\titleformat{\section}{\normalsize\bfseries}{\thesection.}{0.5em}{}
\titleformat{\subsection}{\small\bfseries}{\thesubsection}{0.5em}{}
\titleformat{\subsubsection}{\small\itshape}{}{0em}{}
\titlespacing*{\section}{0pt}{0.8em}{0.25em}
\titlespacing*{\subsection}{0pt}{0.5em}{0.15em}
\titlespacing*{\subsubsection}{0pt}{0.35em}{0.1em}

% ── Compact lists ─────────────────────────────────────────────────────
\setlist[itemize]{leftmargin=1.2em,itemsep=1pt,parsep=0pt,topsep=2pt}
\setlist[enumerate]{leftmargin=1.4em,itemsep=1pt,parsep=0pt,topsep=2pt}

% ── Paragraph spacing ────────────────────────────────────────────────
\setlength{\parskip}{3pt plus 1pt}
\setlength{\parindent}{1em}

% ══════════════════════════════════════════════════════════════════════
\begin{document}

% ── Title ─────────────────────────────────────────────────────────────
\twocolumn[{%
\centering
{\Large\bfseries DeltaScribe Edge: Offline Longitudinal Chest X-Ray Copilot\\[2pt]
with Verifiable Evidence and FHIR Export\par}
\vspace{6pt}
{\normalsize The MedGemma Impact Challenge --- Kaggle 2026\par}
\vspace{10pt}
\rule{0.4\textwidth}{0.4pt}
\vspace{8pt}
}]
\thispagestyle{empty}

% ══════════════════════════════════════════════════════════════════════
\section{Project Name}

DeltaScribe Edge --- Offline longitudinal chest X-ray copilot with verifiable evidence and FHIR export.

% ══════════════════════════════════════════════════════════════════════
\section{Your Team}

\textbf{Simon} --- Architecture, implementation, and evaluation (independent, end-to-end development).

% ══════════════════════════════════════════════════════════════════════
\section{Problem Statement}

\subsection{Problem Domain}

Follow-up respiratory imaging is among the most common and time-critical workflows in clinical care (pneumonia follow-up, pulmonary edema monitoring, post-treatment assessment).
The core challenge is not interpreting a single image but answering: \emph{what changed since the prior study, and how should that change be documented consistently?}

Clinicians must locate the relevant prior, compare findings across studies, draft standardised interval-change language, and communicate uncertainty.
This process is structured and repetitive yet remains largely manual.
Many clinical environments cannot rely on large closed cloud models due to privacy constraints, limited connectivity, latency, and governance requirements.
Open-weight healthcare models enable local deployment with auditable behaviour, making AI-assisted longitudinal documentation feasible.

Today, the dominant workflow involves the radiologist opening both studies side-by-side, mentally cataloguing findings, and free-text dictating a report.
Variability across reporters is well-documented: identical imaging pairs can yield materially different interval-change assessments depending on time pressure, template availability, and individual reporting style.
A tool that produces a structured, evidence-linked first draft---subject to clinician review---can reduce this variability while preserving clinical judgment.

\subsection{Impact Potential}

An effective solution improves:
(1)~documentation efficiency via faster time-to-first-draft;
(2)~consistency through standardised delta language;
(3)~continuity of care for rotating staff; and
(4)~trustworthiness via explicit evidence mapping and uncertainty gating.

\smallskip
\noindent\textbf{Illustrative estimate.}
If a site drafts 30 follow-up CXR reports per day and saves 2~minutes per report, that yields ${\sim}60$~min/day (${\sim}5$~hours/week) of clinician time returned to direct care.
Across a 10-site deployment, this scales to ${\sim}50$~hours/week of recovered clinician capacity.
With an average time-to-first-draft of 177\,s on commodity hardware (Apple~M3, CPU-only Docker), a 2-minute saving per report is achievable on GPU-equipped workstations.

% ══════════════════════════════════════════════════════════════════════
\section{Overall Solution}

DeltaScribe Edge is an offline-first demonstration application for longitudinal CXR follow-ups.
It accepts:

\begin{itemize}
  \item Current CXR + Prior CXR (required).
  \item Optional: dictation audio (brief clinician note).
  \item Optional: FHIR-formatted patient context (conditions, medications, prior report snippet).
  \item Optional: retrieval evidence (similar historical cases, guideline/template snippets).
\end{itemize}

\noindent It produces four outputs:

\begin{enumerate}
  \item \textbf{Delta Summary} --- concise interval-change bullets.
  \item \textbf{Draft Report} --- radiology-style Findings / Impression template.
  \item \textbf{Evidence Ledger} --- every material claim linked to evidence sources and an uncertainty status.
  \item \textbf{FHIR Export} --- a conformant Bundle containing \texttt{DiagnosticReport} and \texttt{Observation} resources, generated by deterministic mapping from the structured output (the model does not write FHIR directly).
\end{enumerate}

\subsection{Why HAI-DEF / MedGemma}

MedGemma is purpose-built for medical multimodal reasoning, including CXR interpretation and longitudinal comparison.
Using open-weight healthcare models also enables full control over weights and infrastructure, supporting offline, privacy-preserving deployment where care is delivered.

\smallskip
\noindent\textbf{HAI-DEF models used:}

\begin{itemize}
  \item \textbf{MedGemma 4B-IT} (text): longitudinal CXR comparison, report drafting, structured JSON output.
  Default deployment uses a GGUF-quantised variant (Q4\_K\_M, ${\sim}$2.3\,GB) via \texttt{llama-cpp-python} for CPU/edge inference; a multimodal transformers pathway is available for GPU environments.
  \item \textbf{MedASR} (optional): offline medical dictation (conditional fallback chain).
  \item \textbf{MedSigLIP / CXR Foundation} (optional): embedding-based retrieval for explainability and evidence grounding.
\end{itemize}

% ══════════════════════════════════════════════════════════════════════
\section{Technical Details}

\subsection{Architecture}

A reproducible containerised stack:
\textbf{Frontend:} Next.js + shadcn/ui (local web UI).
\textbf{Backend:} Python FastAPI orchestrator and inference service.
\textbf{Runtime:} local model weights mounted into the container; no network access required.
\textbf{Assets:} curated demo patient pairs, reference atlas, guideline/template pack, local FHIR bundles.
All inference runs locally; no patient data needs to leave the device.

% ── Architecture Diagram ─────────────────────────────────────────────
\begin{figure}[h]
\centering
\resizebox{\columnwidth}{!}{%
\begin{tikzpicture}[
    node distance=0.35cm and 0.4cm,
    every node/.style={font=\footnotesize},
    box/.style={draw, rounded corners=2pt, minimum height=0.55cm,
                minimum width=2.2cm, align=center, thick},
    io/.style={box, densely dashed},
    arrow/.style={-{Stealth[length=2mm]}, thick},
  ]
  % Inputs
  \node[io] (cxr) {Prior + Current CXR};
  \node[io, below=of cxr] (audio) {Dictation (opt.)};
  \node[io, below=of audio] (fhir_in) {FHIR Context (opt.)};

  % Pipeline
  \node[box, right=1.0cm of audio] (val) {1. Validate};
  \node[box, right=of val] (pre) {2. Preprocess};
  \node[box, below=of pre] (ctx) {3. Context\\Extract};
  \node[box, below=of ctx] (rag) {4. RAG\\Retrieval};
  \node[box, right=of pre] (inf) {5. MedGemma\\Inference};
  \node[box, below=of inf] (schema) {6. Schema\\Validate};
  \node[box, below=of schema] (ledger) {7. Evidence\\Ledger};

  % Outputs
  \node[io, right=1.0cm of inf] (report) {Draft Report};
  \node[io, below=of report] (delta) {Delta Summary};
  \node[io, below=of delta] (ev) {Evidence Ledger};
  \node[io, below=of ev] (fhir_out) {FHIR Export};

  % Arrows — inputs
  \draw[arrow] (cxr.east) -- ++(0.3,0) |- (val.west);
  \draw[arrow] (audio.east) -- (val.west);
  \draw[arrow] (fhir_in.east) -- ++(0.3,0) |- (val.west);

  % Arrows — pipeline
  \draw[arrow] (val) -- (pre);
  \draw[arrow] (pre) -- (inf);
  \draw[arrow] (pre.south) -- (ctx.north);
  \draw[arrow] (ctx.south) -- (rag.north);
  \draw[arrow] (rag.east) -| ($(inf.south west)+(0.3,0)$);
  \draw[arrow] (inf) -- (schema);
  \draw[arrow] (schema) -- (ledger);

  % Arrows — outputs
  \draw[arrow] (inf.east) -- (report.west);
  \draw[arrow] (schema.east) -| ($(delta.west)-(0.2,0)$) -- (delta.west);
  \draw[arrow] (ledger.east) -| ($(ev.west)-(0.2,0)$) -- (ev.west);
  \draw[arrow] (ledger.south) -| (fhir_out.west);

\end{tikzpicture}%
}
\caption{DeltaScribe Edge agentic pipeline. Dashed boxes are inputs/outputs; solid boxes are deterministic tool steps. Each step emits a visible event to the UI timeline.}
\label{fig:pipeline}
\end{figure}

\subsection{Agentic Workflow}

DeltaScribe Edge decomposes the clinical workflow into a \emph{deterministic state machine} with explicit tool steps (Figure~\ref{fig:pipeline}), each emitting a visible event in the UI timeline:

\begin{enumerate}
  \item \textbf{Intake} --- verify prior/current imaging present; check optional inputs.
  \item \textbf{Preprocessing} --- normalise images to consistent internal format.
  \item \textbf{Context retrieval} (optional) --- parse minimal fields from local FHIR bundle; transcribe dictation.
  \item \textbf{Evidence retrieval} (optional) --- top-$K$ image RAG from local atlas; text RAG from guideline pack.
  \item \textbf{Longitudinal inference} --- MedGemma generates delta findings, report draft, and structured JSON.
  \item \textbf{Schema validation and repair} --- enforce strict JSON schema; repair structure only if needed.
  \item \textbf{Assembly} --- build Evidence Ledger (claims $\rightarrow$ evidence pointers $+$ uncertainty), draft report, and FHIR Bundle (\texttt{DiagnosticReport} + \texttt{Observations}).
\end{enumerate}

This design matches the Agentic Workflow Prize intent: a significant overhaul of a complex process using callable tools with visible intermediate artifacts and validation gates.

\subsection{Evidence Ledger (Verifiable Behavior)}

Every key statement is represented as a \emph{claim object}: (i)~claim text, (ii)~evidence pointers (current image, prior image, retrieved cases, guideline snippets, dictation excerpts), (iii)~uncertainty level $\in \{\textit{low},\, \textit{medium},\, \textit{high}\}$, and (iv)~clinician-readable rationale.

\smallskip
\noindent\textbf{Policy: no evidence $\Rightarrow$ no valid finding.}
The schema requires every finding to link at least one evidence source; findings without evidence cannot pass validation.
This structural constraint ensures that no unsupported claim reaches the final output.

% ── Evidence Ledger Example ──────────────────────────────────────────
\begin{table}[h]
\centering
\small
\caption{Example Evidence Ledger entry for a single claim.}
\label{tab:ledger_example}
\begin{tabularx}{\columnwidth}{@{}lX@{}}
\toprule
\textbf{Field} & \textbf{Value} \\
\midrule
Claim       & Right lower lobe opacity improved vs.\ prior. \\
Evidence    & Current CXR (region 4), Prior CXR (region 4), Atlas case \#17 (sim.\ 0.82). \\
Uncertainty & \textit{low} \\
Rationale   & Opacity extent decreased; consistent with retrieved atlas exemplar showing resolving consolidation. \\
\bottomrule
\end{tabularx}
\end{table}

\subsection{Retrieval Augmentation}

\textbf{Image RAG.}
BiomedCLIP (Microsoft, via \texttt{open\_clip}) encodes the query CXR; the top-$K$ most similar cases are retrieved from a precomputed atlas of CXR Foundation embeddings via cosine similarity.
Retrieved exemplars are displayed with similarity scores.

\textbf{Zero-shot classification \& Text RAG.}
MedSigLIP provides zero-shot pathology classification and cross-modal guideline retrieval from an offline pack of 15 clinical guidelines.
Both retrieval pathways are visible to the clinician in the UI, supporting transparency and enabling informed override.

\subsection{Key Design Decisions}

Several architectural choices distinguish DeltaScribe Edge from a naive ``model-generates-report'' approach:

\begin{enumerate}
  \item \textbf{Deterministic pipeline, not end-to-end generation.}
  Each step is independently testable, debuggable, and auditable.
  A failure at step~6 (schema validation) does not require re-running inference.

  \item \textbf{Structured intermediate representation.}
  MedGemma outputs a constrained JSON schema per finding: \texttt{label} (8-class vocabulary), \texttt{delta} (improved/stable/worsened/uncertain), \texttt{rationale}, \texttt{evidence\_refs}, optional \texttt{bounding\_box}, and \texttt{uncertainty} level.
  The draft report is rendered \emph{from} this structure, not generated as free text---ensuring consistency between the summary, report, and ledger.

  \item \textbf{Retrieval as context and transparency.}
  RAG is not used to ``improve generation quality'' in the usual sense.
  Retrieved similar cases and guideline snippets are injected into the inference prompt, giving MedGemma grounded context.
  Retrieval artifacts are also surfaced directly in the UI so clinicians can independently verify model output against evidence.

  \item \textbf{FHIR export as a first-class output.}
  Rather than treating interoperability as an afterthought, the FHIR Bundle is generated deterministically from the structured JSON at every run, ensuring the system is integration-ready from the start.
\end{enumerate}

\subsection{Evaluation}

Table~\ref{tab:eval} summarises the evaluation framework.
Metrics are grouped by reliability, feasibility, consistency, and trust to map directly to the judging criteria.

\begin{table}[h]
\centering
\small
\caption{Evaluation metrics and results.}
\label{tab:eval}
\begin{tabularx}{\columnwidth}{@{}lXr@{}}
\toprule
\textbf{Category} & \textbf{Metric} & \textbf{Result} \\
\midrule
Reliability & Schema validity (\% valid JSON w/o repair) & 100\,\% \\
            & FHIR export integrity (structural)          & Pass \\
Feasibility & Latency (time-to-first-draft, M3/CPU)      & 177\,s avg \\
            & Latency range (5 demo pairs)                & 130--231\,s \\
Consistency & Delta label agreement (curated pairs)       & \textit{pending} \\
Trust       & Evidence coverage rate                      & 100\,\% \\
            & Disclaimer present                          & 100\,\% \\
\bottomrule
\end{tabularx}
\end{table}

\noindent\textbf{Evaluation protocol.}
Metrics are computed over 5 curated prior/current CXR demo pairs run end-to-end in Docker on an Apple~M3 (24\,GB RAM, CPU-only, GGUF Q4\_K\_M inference).
Schema validity: 5/5 cases returned valid JSON on first pass (no repair needed).
Latency: mean 177\,s (range 130--231\,s); cold model load (GGUF + retrieval models) accounts for ${\sim}$60\,s of the first run.
Evidence coverage: all asserted findings include $\geq$1 valid evidence pointer.
A disclaimer (``Draft only---requires clinician verification'') is present in 100\% of outputs.
Delta label agreement against reference annotations is planned but not yet computed.
All evaluation code, data, and commands to reproduce results are provided in the public repository.

\subsection{Responsible Use and Limitations}

DeltaScribe Edge is \emph{not} a diagnostic device.
Outputs are drafts intended for review by qualified clinicians.
Longitudinal claims are suppressed when priors are missing; retrieval is used for transparency and grounding---not for ``diagnosis by similarity.''
All inference runs locally; no PHI needs to leave the device.

\smallskip
\noindent\textbf{Known limitations:}

\begin{itemize}
  \item Performance is bounded by MedGemma's pre-training distribution; rare pathologies may be under-represented.
  \item The curated atlas is small and intended for demonstration, not production coverage.
  \item The current UI is a prototype suitable for evaluation, not clinical deployment.
  \item The system handles single-modality (CXR) longitudinal comparison only; extension to CT or multi-modality workflows is future work.
\end{itemize}

\subsection{Future Work}

Several directions are natural extensions of the current system:

\begin{itemize}
  \item \textbf{Fine-tuning for longitudinal tasks.}
  Adapting MedGemma with curated prior/current CXR pairs and clinician-annotated interval-change labels to improve delta accuracy beyond zero-shot performance.
  \item \textbf{Expanded atlas and retrieval.}
  Scaling the curated atlas from demonstration size to a clinically representative collection, with stratified sampling across pathology types and severity.
  \item \textbf{Multi-modality support.}
  Extending the pipeline to handle CT follow-ups and cross-modality comparisons (e.g., CXR $\rightarrow$ CT progression tracking).
  \item \textbf{Edge deployment (implemented).}
  MedGemma 4B-IT is quantised to GGUF Q4\_K\_M (${\sim}$2.3\,GB) and runs entirely on CPU via \texttt{llama-cpp-python} inside Docker.
  Evaluated on Apple~M3 (24\,GB) with an average latency of 177\,s per case across 5~demo pairs---no GPU required.
  \item \textbf{Clinical pilot.}
  Partnering with a clinical site to measure real-world time savings and reporting consistency in a controlled pilot study, with IRB approval.
\end{itemize}

\subsection{What the Video Demonstrates}

The 3-minute video shows the end-to-end workflow:

\begin{itemize}
  \item Offline proof (no network access during inference).
  \item Loading a prior/current CXR pair into the local UI.
  \item Agent timeline with visible step-by-step execution and intermediate artifacts.
  \item Evidence Ledger inspection (claim $\rightarrow$ source tracing, as in Table~\ref{tab:ledger_example}).
  \item FHIR export generation and bundle validation.
\end{itemize}

% ══════════════════════════════════════════════════════════════════════
\section{Links}
\label{sec:links}

\noindent\textbf{Track selection:} Main Track + Agentic Workflow Prize.

\smallskip
\noindent\textbf{Required:}
\begin{itemize}
  \item Video demo ($\leq$3\,min): \texttt{[PASTE\_LINK]}
  \item Public code repository: \texttt{[PASTE\_LINK]}
\end{itemize}

\noindent\textbf{Bonus:}
\begin{itemize}
  \item Live demo app: \texttt{[PASTE\_LINK]}
  \item Model artifacts (HAI-DEF lineage): \texttt{[PASTE\_LINK]}
  \item Dataset / atlas curation notes: \texttt{[PASTE\_LINK]}
\end{itemize}

% ══════════════════════════════════════════════════════════════════════
\bibliographystyle{plain}
\begin{thebibliography}{1}
\small

\bibitem{medgemma2026}
F.~Mahvar, Y.~Liu, D.~Golden, F.~Jamil, S.~Jansen, C.~Kirmizi, R.~Pilgrim, D.\,F.~Steiner, A.~Sellergren, R.~Tiwari, S.~Virmani, L.~Yatziv, R.~Hemenway, Y.~Matias, R.\,Levavi~Morad, A.~Hassidim, S.~Shetty, and M.~Cruz.
\newblock The {MedGemma} Impact Challenge.
\newblock \url{https://kaggle.com/competitions/med-gemma-impact-challenge}, 2026.

\end{thebibliography}

\end{document}
